---
title: "Brownlows Bets"
author: "Your Name"
date: "Year 2020"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
```

<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

```{r}
URN = 1:365
```

# What are the fair odds  

```{r}
classSize = 35
winAmount=1
loseAmount=4.25

test = vector('numeric')
saveBank=vector('numeric')
for (j in 1:30){
  bank=0
for(i in 1:637){
test = sample(URN, size=classSize, replace=TRUE)
ifelse(length(unique(test))==classSize, (bank=bank-loseAmount),(bank=bank+winAmount))
saveBank[j] = bank
}
}
hist(saveBank)

```

```{r}
favstats(saveBank)
```
An odds ratio of 4.25:1 seems fair because it returns a low mean which means that it does not heavily favor one side or the other.

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->


# What size of class makes the 2:1 bet fair?  
<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->

```{r}
URN = 1:365
```

```{r}
classSize = 29
winAmount=1
loseAmount=2

test = vector('numeric')
saveBank=vector('numeric')
for (j in 1:30){
  bank=0
for(i in 1:637){
test = sample(URN, size=classSize, replace=TRUE)
ifelse(length(unique(test))==classSize, (bank=bank-loseAmount),(bank=bank+winAmount))
saveBank[j] = bank
}
}
hist(saveBank)

```

```{r}
favstats(saveBank)
```

A class size between 28 and 29 people would make for a fair bet. At 28 people the odds are in favor of the class and at 29 they're in favor of the professor.

# In the class and payoff as described what is a 95% confidence interval for Brownlow's earnings?
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->

```{r}
classSize = 35
winAmount=1
loseAmount=2.00

test = vector('numeric')
saveBank=vector('numeric')
for (j in 1:1000){
  bank=0
for(i in 1:637){
test = sample(URN, size=classSize, replace=TRUE)
ifelse(length(unique(test))==classSize, (bank=bank-loseAmount),(bank=bank+winAmount))
saveBank[j] = bank
}
}
hist(saveBank)

```

```{r}
sortedBank <- sort(saveBank)
lcl<-sortedBank[25]
ucl<-sortedBank[975]
print(paste(lcl,ucl))
```

The 95% confidence would be between $220 and $334.  He would make (on average) a minimum of $220 and a maximum of $334. We found this by finding the middle 95% (above 25 and below 950 for 1000 bets).

# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

He should increase his payout if his class size is around 35 students because once he goes over 28 students, the odds are in his favor. However, if his class size is 28 or less the odds are against him and he would want to decrease his payout in order to have his profits center on zero. The ideal class size for running this bet while issuing a 2:1 payout would be between 28 and 29 students if the goal was to have a profit centered on zero.
